### 2022-11-25 MSDS 6371 Project - Yule and Kondakindi.qmd
Initial commit

---
title: "MSDS 6371 Project - Fall 2022"
author: "Andrew Yule, Krithika Kondakindi"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console
---

# Introduction

# Data Description

# Analysis

### Question 1

Brief introduction to the questions of interest and the setting of the problem.

#### Restatement of the Problem

(Where did the data come from? How big is it? How many observations? Where can we find out more? What are the specific variables that we need to know with respect to your analysis?)

#### Build and Fit the Model

#### Checking the Assumptions

Residual Plots Influential point analysis (Cook's D and Leverage) Make sure to address each assumption.

#### Comparing Competing Models

Adj R2 Internal CV Press

#### Parameters

Estimates Interpretation Confidence Intervals

#### Conclusion

A short summary of the analysis.

### R Shiny: Price v. Living Area Chart

### Question 2

#### Restatement of Problem

#### Model Selection

Type of Selection Stepwise Forward Backward Custom

#### Checking Assumptions

Residual Plots Influential point analysis (Cook's D and Leverage) Make sure to address each assumption

#### Comparing Competing Models

Adj R2 Internal CV Press\
Kaggle Score

#### Conclusion

A short summary of the analysis

# Appendix

```{r}
library(tidyverse)
```

Read training and test data and perform any cleaning steps
```{r}
housesTrain = read_csv("train.csv")
housesTest = read_csv("test.csv")

remove(houses)
```

### 2022-12-01 Analysis 2.qmd
Analysis 2: Moved code from main markdown file into separate one to focus on Analysis 2 question

---
title: "Analysis 2"
author: "Andrew Yule"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console
---

Load required libraries

```{r}
library(tidyverse)
library(broom)
library(terra)
```

Read training and test data and perform any cleaning steps

```{r}
housesTrain = read_csv("train.csv")
housesTest = read_csv("test.csv")

# Remove ID column only from the training set
housesTrain = select(housesTrain, -Id)

# Find which columns have missing values between the training and test data sets
colsToDrop = housesTrain |>
  select_if(function(x) any(is.na(x))) |>
  colnames()
colsToDrop2 = housesTest |>
  select_if(function(x) any(is.na(x))) |>
  colnames()

# Remove any of the columns that had missing values
housesTrain =  housesTrain |>
  select(-starts_with(c(colsToDrop, colsToDrop2)))
housesTest =  housesTest |>
  select(-starts_with(c(colsToDrop, colsToDrop2)))

hist(housesTrain$LotArea)

# From the plot below, LotArea should likely cut any value off at 30,000 sq ft
housesTrain$LotArea = clamp(housesTrain$LotArea, lower = 0, upper = 30000)
housesTest$LotArea = clamp(housesTest$LotArea, lower = 0, upper = 30000)

# From the plot below, GrLivArea should likely cut any value off at 4000 sq ft
housesTrain$GrLivArea = clamp(housesTrain$GrLivArea, lower = 0, upper = 4000)
housesTest$GrLivArea = clamp(housesTest$GrLivArea, lower = 0, upper = 4000)

# From the plot below, MSSubClass looks to have a quadratic relationship

# From the plot below, LotArea should should have a log relationship, which means each value should be greater than or equal to 1
housesTrain$LotArea = clamp(housesTrain$LotArea, lower = 1)
housesTest$LotArea = clamp(housesTest$LotArea, lower = 1)

# From the plot below, OpenPorchSF should should have a log relationship, which means each value should be greater than or equal to 1
#housesTrain$OpenPorchSF = clamp(housesTrain$OpenPorchSF, lower = 1)
#housesTest$OpenPorchSF = clamp(housesTest$OpenPorchSF, lower = 1)

# From the plot below, MiscVal should should have a log relationship, which means each value should be greater than or equal to 1
housesTrain$MiscVal = clamp(housesTrain$MiscVal, lower = 1)
housesTest$MiscVal = clamp(housesTest$MiscVal, lower = 1)

# From the plot below, PoolArea should should have a log relationship, which means each value should be greater than or equal to 1
#housesTrain$PoolArea = clamp(housesTrain$PoolArea, lower = 1)
#housesTest$PoolArea = clamp(housesTest$PoolArea, lower = 1)

# Screen for outliers by looking at the histograms of any variables that are all numeric
housesTrain |> 
  select(where(is.numeric)) |>
  gather() |>
  ggplot(aes(x = value)) + 
    facet_wrap(~key, scales = "free") + 
    geom_histogram()

# MoSold should be a factor instead of integer
housesTrain$MoSold = factor(housesTrain$MoSold, levels = seq(1, 12))
housesTest$MoSold = factor(housesTest$MoSold, levels = seq(1, 12))

# The sale price has a wide variance. Applying a log transformation helps
housesTrain |>
  ggplot(aes(x = log(SalePrice))) +
  geom_histogram()

# Log of SalePrice should be used to reduce the variance
housesTrain$SalePrice = log(housesTrain$SalePrice)

```

Full model creation

```{r}
housesTrain |> count(`3SsnPorch`)

fit = lm(SalePrice ~ I(MSSubClass^2) + log(LotArea) + Street + LotShape + LandContour + LotConfig + LandSlope + Neighborhood + Condition1 + Condition2 + BldgType + HouseStyle + OverallQual + OverallCond + YearBuilt + YearRemodAdd + RoofStyle + RoofMatl + ExterQual + ExterCond + Foundation + Heating + HeatingQC + CentralAir + `1stFlrSF` + `2ndFlrSF` + LowQualFinSF + GrLivArea + FullBath + HalfBath + BedroomAbvGr + KitchenAbvGr + TotRmsAbvGrd + Fireplaces + PavedDrive + WoodDeckSF + OpenPorchSF + EnclosedPorch + `3SsnPorch` + ScreenPorch + PoolArea + log(MiscVal) + MoSold + YrSold + SaleCondition, data = housesTrain)
summary(fit)

```

Test the basic model against the test data set and submit to Kaggle to see the score

```{r}
results = housesTest
results$SalePrice = exp(predict.lm(fit, newdata = housesTest))
results = results |>
  select(Id, SalePrice)

# Export
write_csv(results, "test_submission_10.csv")
```

Try using forward stepping approach

```{r}
# Define intercept-only model, this is also just the median sale price
intercept_only = lm(SalePrice ~ 1, data = housesTrain)

# Define model with all predictors
all = lm(SalePrice ~ ., data = housesTrain)

# Perform forward step-wise regression
forward = step(intercept_only, direction = 'forward', scope = formula(all), trace = 0)

# View final model
summary(forward)

```

Test the forward model against the test data set and submit to Kaggle

```{r}
results = housesTest
results$SalePrice = exp(predict.lm(forward, newdata = housesTest))
results = results |>
  select(Id, SalePrice)

# Export
write_csv(results, "test_submission_forward_1.csv")
```

Try using backward stepping approach

```{r}
# Define model with all predictors
all = lm(SalePrice ~ ., data = housesTrain)

# Perform forward step-wise regression
backward = step(all, direction = 'backward', scope = formula(all), trace = 0)

# View final model
summary(backward)

```

Test the backward model against the test data set and submit to Kaggle

```{r}
results = housesTest
results$SalePrice = exp(predict.lm(backward, newdata = housesTest))
results = results |>
  select(Id, SalePrice)

# Export
write_csv(results, "test_submission_backward_1.csv")
```

Try using stepwise stepping approach

```{r}
# Define intercept-only model, this is also just the average fat amount
intercept_only = lm(SalePrice ~ 1, data = housesTrain)

# Define model with all predictors
all = lm(SalePrice ~ ., data = housesTrain)

# Perform forward step-wise regression
both = step(intercept_only, direction = 'both', scope = formula(all), trace = 0)

# View final model
summary(both)
```

Test the stepwise model against the test data set and submit to Kaggle

```{r}
results = housesTest
results$SalePrice = exp(predict.lm(both, newdata = housesTest))
results = results |>
  select(Id, SalePrice)

# Export
write_csv(results, "test_submission_stepwise_1.csv")
```

### 2022-12-04 Analysis 2.qmd
Updates

@@
 ```{r}
+#library(MASS)
+library(olsrr)
 library(tidyverse)
+library(terra)
+
 library(broom)
-library(terra)
+library(caret)
 ```
 
-Read training and test data and perform any cleaning steps
+Read training and test data, then drop any columns containing missing values
 
@@
 # Remove ID column only from the training set
-housesTrain = select(housesTrain, -Id)
+housesTrain = dplyr::select(housesTrain, -Id)
 
@@
 housesTrain =  housesTrain |>
-  select(-starts_with(c(colsToDrop, colsToDrop2)))
+  dplyr::select(-starts_with(c(colsToDrop, colsToDrop2)))
 housesTest =  housesTest |>
-  select(-starts_with(c(colsToDrop, colsToDrop2)))
+  dplyr::select(-starts_with(c(colsToDrop, colsToDrop2)))
+```
 
-hist(housesTrain$LotArea)
-
-# From the plot below, LotArea should likely cut any value off at 30,000 sq ft
-housesTrain$LotArea = clamp(housesTrain$LotArea, lower = 0, upper = 30000)
-housesTest$LotArea = clamp(housesTest$LotArea, lower = 0, upper = 30000)
-
-# From the plot below, GrLivArea should likely cut any value off at 4000 sq ft
-housesTrain$GrLivArea = clamp(housesTrain$GrLivArea, lower = 0, upper = 4000)
-housesTest$GrLivArea = clamp(housesTest$GrLivArea, lower = 0, upper = 4000)
-
-# From the plot below, MSSubClass looks to have a quadratic relationship
-
-# From the plot below, LotArea should should have a log relationship, which means each value should be greater than or equal to 1
-housesTrain$LotArea = clamp(housesTrain$LotArea, lower = 1)
-housesTest$LotArea = clamp(housesTest$LotArea, lower = 1)
-
-# From the plot below, OpenPorchSF should should have a log relationship, which means each value should be greater than or equal to 1
-#housesTrain$OpenPorchSF = clamp(housesTrain$OpenPorchSF, lower = 1)
-#housesTest$OpenPorchSF = clamp(housesTest$OpenPorchSF, lower = 1)
-
-# From the plot below, MiscVal should should have a log relationship, which means each value should be greater than or equal to 1
-housesTrain$MiscVal = clamp(housesTrain$MiscVal, lower = 1)
-housesTest$MiscVal = clamp(housesTest$MiscVal, lower = 1)
-
-# From the plot below, PoolArea should should have a log relationship, which means each value should be greater than or equal to 1
-#housesTrain$PoolArea = clamp(housesTrain$PoolArea, lower = 1)
-#housesTest$PoolArea = clamp(housesTest$PoolArea, lower = 1)
-
-# Screen for outliers by looking at the histograms of any variables that are all numeric
+```{r}
+# Screen for numerical variables that should be log-transformed
 housesTrain |> 
-  select(where(is.numeric)) |>
+  dplyr::select(where(is.numeric)) |>
   gather() |>
@@
     geom_histogram()
+```
+
+The following variables should be log transformed:
+- LotArea
+- FirstFlrSF
+- SecondFlrSF
+- GrLivArea
+- WoodDeckSF
+- OpenPorchSF
+- EnclosedPorch
+- ThirdSsnPorch
+- ScreenPorch
+- PoolArea
+- MiscVal
+- SalePrice
+
+```{r}
 
@@
 
-# The sale price has a wide variance. Applying a log transformation helps
-housesTrain |>
-  ggplot(aes(x = log(SalePrice))) +
-  geom_histogram()
+# First ensure that any variables that need log transformations are set to a minimum value of 1 instead of 0. Then, take the log of the values.
+housesTrain = housesTrain |>
+  mutate(
+    LotArea = log(clamp(LotArea, lower = 1)),
+    FirstFlrSF = log(clamp(`1stFlrSF`, lower = 1)),
+    SecondFlrSF = log(clamp(`2ndFlrSF`, lower = 1)),
+    GrLivArea = log(clamp(GrLivArea, lower = 1)),
+    WoodDeckSF = log(clamp(WoodDeckSF, lower = 1)),
+    OpenPorchSF = log(clamp(OpenPorchSF, lower = 1)),
+    EnclosedPorch = log(clamp(EnclosedPorch, lower = 1)),
+    ThirdSsnPorch = log(clamp(`3SsnPorch`, lower = 1)),
+    ScreenPorch = log(clamp(ScreenPorch, lower = 1)),
+    PoolArea = log(clamp(PoolArea, lower = 1)),
+    MiscVal = log(clamp(MiscVal, lower = 1)),
+    SalePrice = log(clamp(SalePrice, lower = 1)),
+  )
 
-# Log of SalePrice should be used to reduce the variance
-housesTrain$SalePrice = log(housesTrain$SalePrice)
+housesTest = housesTest |>
+  mutate(
+    LotArea = log(clamp(LotArea, lower = 1)),
+    FirstFlrSF = log(clamp(`1stFlrSF`, lower = 1)),
+    SecondFlrSF = log(clamp(`2ndFlrSF`, lower = 1)),
+    GrLivArea = log(clamp(GrLivArea, lower = 1)),
+    WoodDeckSF = log(clamp(WoodDeckSF, lower = 1)),
+    OpenPorchSF = log(clamp(OpenPorchSF, lower = 1)),
+    EnclosedPorch = log(clamp(EnclosedPorch, lower = 1)),
+    ThirdSsnPorch = log(clamp(`3SsnPorch`, lower = 1)),
+    ScreenPorch = log(clamp(ScreenPorch, lower = 1)),
+    PoolArea = log(clamp(PoolArea, lower = 1)),
+    MiscVal = log(clamp(MiscVal, lower = 1))
+  )
+```
+
+Forward step regression
+```{r}
+forwardResults = ols_step_forward_aic(lm(SalePrice ~ ., data = housesTrain))
+forwardModel = forwardResults$model
+
+plot(forwardResults)
+
+summary(forwardModel)
+
+results = housesTest
+results$SalePrice = exp(predict.lm(forwardModel, newdata = housesTest))
+results = results |>
+  dplyr::select(Id, SalePrice)
+
+# Export
+write_csv(results, "Submissions/Forward_Model.csv")
 
@@
 
-Full model creation
+Backward step regression
+```{r}
+backwardResults = ols_step_backward_aic(lm(SalePrice ~ ., data = housesTrain))
+backwardModel = backwardResults$model
 
-```{r}
-housesTrain |> count(`3SsnPorch`)
+plot(backwardResults)
 
-fit = lm(SalePrice ~ I(MSSubClass^2) + log(LotArea) + Street + LotShape + LandContour + LotConfig + LandSlope + Neighborhood + Condition1 + Condition2 + BldgType + HouseStyle + OverallQual + OverallCond + YearBuilt + YearRemodAdd + RoofStyle + RoofMatl + ExterQual + ExterCond + Foundation + Heating + HeatingQC + CentralAir + `1stFlrSF` + `2ndFlrSF` + LowQualFinSF + GrLivArea + FullBath + HalfBath + BedroomAbvGr + KitchenAbvGr + TotRmsAbvGrd + Fireplaces + PavedDrive + WoodDeckSF + OpenPorchSF + EnclosedPorch + `3SsnPorch` + ScreenPorch + PoolArea + log(MiscVal) + MoSold + YrSold + SaleCondition, data = housesTrain)
-summary(fit)
+summary(backwardModel)
+
+results = housesTest
+results$SalePrice = exp(predict.lm(backwardModel, newdata = housesTest))
+results = results |>
+  dplyr::select(Id, SalePrice)
+
+# Export
+write_csv(results, "Submissions/Backward_Model.csv")
 
@@
 
-Test the basic model against the test data set and submit to Kaggle to see the score
+Stepwise regression
+```{r}
+stepResults = ols_step_both_aic(lm(SalePrice ~ ., data = housesTrain), details = TRUE)
+stepModel = lm(SalePrice ~ OverallQual + GrLivArea + Neighborhood + OverallCond + YearBuilt + LotArea + RoofMatl + KitchenAbvGr + SaleCondition + Condition2 + Foundation + Fireplaces + Heating + ExterQual + Condition1 + PoolArea + ScreenPorch + WoodDeckSF + HeatingQC + CentralAir + BedroomAbvGr + FirstFlrSF + `2ndFlrSF` + Street + LandSlope + HalfBath + EnclosedPorch + SecondFlrSF + MiscVal + PavedDrive + BldgType + `1stFlrSF` + ExterCond + LotConfig + YearRemodAdd, data = housesTrain)
 
-```{r}
+plot(stepResults)
+
+summary(stepModel)
+
 results = housesTest
-results$SalePrice = exp(predict.lm(fit, newdata = housesTest))
+results$SalePrice = exp(predict.lm(stepModel, newdata = housesTest))
 results = results |>
@@
 # Export
-write_csv(results, "test_submission_10.csv")
-```
-
-Try using forward stepping approach
-
-```{r}
-# Define intercept-only model, this is also just the median sale price
-intercept_only = lm(SalePrice ~ 1, data = housesTrain)
-
-# Define model with all predictors
-all = lm(SalePrice ~ ., data = housesTrain)
-
-# Perform forward step-wise regression
-forward = step(intercept_only, direction = 'forward', scope = formula(all), trace = 0)
-
-# View final model
-summary(forward)
+write_csv(results, "Submissions/Both_Model.csv")
 
@@
 
-Test the forward model against the test data set and submit to Kaggle
+Best custom model using caret
+```{r}
+customModel = lm(SalePrice ~ I(MSSubClass^2) + log(LotArea) + Street + LotShape + LandContour + LotConfig + LandSlope + Neighborhood + Condition1 + Condition2 + BldgType + HouseStyle + OverallQual + OverallCond + YearBuilt + YearRemodAdd + RoofStyle + RoofMatl + ExterQual + ExterCond + Foundation + Heating + HeatingQC + CentralAir + `1stFlrSF` + `2ndFlrSF` + LowQualFinSF + GrLivArea + FullBath + HalfBath + BedroomAbvGr + KitchenAbvGr + TotRmsAbvGrd + Fireplaces + PavedDrive + WoodDeckSF + OpenPorchSF + EnclosedPorch + `3SsnPorch` + ScreenPorch + PoolArea + log(MiscVal) + MoSold + YrSold + SaleCondition, data = housesTrain)
+summary(fit)
 
-```{r}
+#control = trainControl(method = "LOOCV", number = 10)
+control = trainControl(method = "cv")
+customModel = train(SalePrice ~ ., data = housesTrain, method = "lm", trControl = control)
+summary(customModel)
+
 results = housesTest
-results$SalePrice = exp(predict.lm(forward, newdata = housesTest))
+results$SalePrice = exp(predict(customModel, newdata = housesTest))
 results = results |>
@@
 # Export
-write_csv(results, "test_submission_forward_1.csv")
-```
-
-Try using backward stepping approach
-
-```{r}
-# Define model with all predictors
-all = lm(SalePrice ~ ., data = housesTrain)
-
-# Perform forward step-wise regression
-backward = step(all, direction = 'backward', scope = formula(all), trace = 0)
-
-# View final model
-summary(backward)
+write_csv(results, "Submissions/Custom_Model.csv")
 
@@
 
-Test the backward model against the test data set and submit to Kaggle
+Best custom model using external cross validation
+```{r}
+customModel = lm(SalePrice ~ I(MSSubClass^2) + log(LotArea) + Street + LotShape + LandContour + LotConfig + LandSlope + Neighborhood + Condition1 + Condition2 + BldgType + HouseStyle + OverallQual + OverallCond + YearBuilt + YearRemodAdd + RoofStyle + RoofMatl + ExterQual + ExterCond + Foundation + Heating + HeatingQC + CentralAir + `1stFlrSF` + `2ndFlrSF` + LowQualFinSF + GrLivArea + FullBath + HalfBath + BedroomAbvGr + KitchenAbvGr + TotRmsAbvGrd + Fireplaces + PavedDrive + WoodDeckSF + OpenPorchSF + EnclosedPorch + `3SsnPorch` + ScreenPorch + PoolArea + log(MiscVal) + MoSold + YrSold + SaleCondition, data = housesTrain)
+summary(fit)
 
-```{r}
+#control = trainControl(method = "LOOCV", number = 10)
+control = trainControl(method = "cv")
+customModel = train(SalePrice ~ ., data = housesTrain, method = "lm", trControl = control)
+summary(customModel)
+
 results = housesTest
-results$SalePrice = exp(predict.lm(backward, newdata = housesTest))
+results$SalePrice = exp(predict(customModel, newdata = housesTest))
 results = results |>
@@
 # Export
-write_csv(results, "test_submission_backward_1.csv")
+write_csv(results, "Submissions/Custom_Model.csv")
+
 ```
-
-Try using stepwise stepping approach
-
-```{r}
-# Define intercept-only model, this is also just the average fat amount
-intercept_only = lm(SalePrice ~ 1, data = housesTrain)
-
-# Define model with all predictors
-all = lm(SalePrice ~ ., data = housesTrain)
-
-# Perform forward step-wise regression
-both = step(intercept_only, direction = 'both', scope = formula(all), trace = 0)
-
-# View final model
-summary(both)
-```
-
-Test the stepwise model against the test data set and submit to Kaggle
-
-```{r}
-results = housesTest
-results$SalePrice = exp(predict.lm(both, newdata = housesTest))
-results = results |>
-  select(Id, SalePrice)
-
-# Export
-write_csv(results, "test_submission_stepwise_1.csv")
-```

### 2022-12-05 Analysis 2.qmd
Updates to main write up

@@
 Load required libraries
-
 ```{r}
-#library(MASS)
+library(MASS)
 library(olsrr)
+library(caret)
+library(terra)
+library(performance)
 library(tidyverse)
-library(terra)
-
 library(broom)
-library(caret)
 ```
@@
 Read training and test data, then drop any columns containing missing values
-
 ```{r}
@@
 
-The following variables should be log transformed:
+The following variables can be log transformed in order to reduce their variance:
 - LotArea
@@
 
+Additionally, MoSold variable was identified as needing to be a factor instead of an integer
 ```{r}
-
 # MoSold should be a factor instead of integer
@@
 
+# Perform any transformations on the test data as well
 housesTest = housesTest |>
@@
 
-Forward step regression
+Forward step regression using olss
 ```{r}
-forwardResults = ols_step_forward_aic(lm(SalePrice ~ ., data = housesTrain))
+forwardResults = ols_step_forward_aic(lm(SalePrice ~ ., data = housesTrain), details = TRUE, progress = TRUE)
 forwardModel = forwardResults$model
+summary(forwardModel)
+
+# Retrieved from the output
+# forwardModel = lm(SalePrice ~ OverallQual + GrLivArea + Neighborhood + OverallCond + HouseStyle + YearBuilt + LotArea + RoofMatl + KitchenAbvGr + SaleCondition + Condition2 + Foundation + Fireplaces + Heating + ExterQual + Condition1 + PoolArea + ScreenPorch + WoodDeckSF + HeatingQC + CentralAir + BedroomAbvGr + FirstFlrSF + `2ndFlrSF` + Street + LandSlope + HalfBath + EnclosedPorch + SecondFlrSF + MiscVal + PavedDrive + `1stFlrSF` + BldgType + ExterCond + YearRemodAdd, data = housesTrain)
 
@@
 
-summary(forwardModel)
+plot(forwardModel)
+#performance::check_model(forwardModel, check = c("linearity", "normality", "qq"))
 
@@
 write_csv(results, "Submissions/Forward_Model.csv")
-
 ```
@@
 ```{r}
-backwardResults = ols_step_backward_aic(lm(SalePrice ~ ., data = housesTrain))
+backwardResults = ols_step_backward_aic(lm(SalePrice ~ ., data = housesTrain), details = TRUE, progress = TRUE)
 backwardModel = backwardResults$model
+
+# backwardModel = lm(SalePrice ~ LotArea + Street + LotConfig + LandSlope + Neighborhood + Condition1 + Condition2 + BldgType + OverallQual + OverallCond + YearBuilt + YearRemodAdd + RoofMatl + ExterQual + ExterCond + Foundation + Heating + HeatingQC + CentralAir + `1stFlrSF` + `2ndFlrSF` + GrLivArea + HalfBath + BedroomAbvGr + KitchenAbvGr + Fireplaces + PavedDrive + WoodDeckSF + EnclosedPorch + ScreenPorch + PoolArea + MiscVal + SaleCondition + FirstFlrSF + SecondFlrSF, data = housesTrain)
 
@@
 write_csv(results, "Submissions/Backward_Model.csv")
-
 ```
@@
 stepResults = ols_step_both_aic(lm(SalePrice ~ ., data = housesTrain), details = TRUE)
+
+# ols_step_both_aic does not have a direct output for the model determined so we have to write the results as we go and then copy the final model produced and paste it below
 stepModel = lm(SalePrice ~ OverallQual + GrLivArea + Neighborhood + OverallCond + YearBuilt + LotArea + RoofMatl + KitchenAbvGr + SaleCondition + Condition2 + Foundation + Fireplaces + Heating + ExterQual + Condition1 + PoolArea + ScreenPorch + WoodDeckSF + HeatingQC + CentralAir + BedroomAbvGr + FirstFlrSF + `2ndFlrSF` + Street + LandSlope + HalfBath + EnclosedPorch + SecondFlrSF + MiscVal + PavedDrive + BldgType + `1stFlrSF` + ExterCond + LotConfig + YearRemodAdd, data = housesTrain)
@@
 write_csv(results, "Submissions/Both_Model.csv")
+```
+
+# For the custom model use the variables determined from the forward selection and then try and improve the model using CV
+```{r}
+
+housesTrainSub = housesTrain |>
+  select(OverallQual, GrLivArea, Neighborhood, OverallCond, HouseStyle, YearBuilt, LotArea, RoofMatl, KitchenAbvGr, SaleCondition, Condition2, Foundation, Fireplaces, Heating, ExterQual, Condition1, PoolArea, ScreenPorch, WoodDeckSF, HeatingQC, CentralAir, BedroomAbvGr, FirstFlrSF, `2ndFlrSF`, Street, LandSlope, HalfBath, EnclosedPorch, SecondFlrSF, MiscVal, PavedDrive, `1stFlrSF`, BldgType, ExterCond, YearRemodAdd, SalePrice)
+
+# Define the number of subsets (or "folds") to use
+#train_control = trainControl(method = "cv", number = 5)
+train_control = trainControl(method = "LOOCV")
+
+#train the model
+customModel = train(SalePrice ~ ., data = housesTrainSub, method = "lm", trControl = train_control)
+print(customModel)
+summary(customModel)
+
+results = housesTest
+results$SalePrice = exp(predict(customModel, newdata = housesTest))
+results = results |>
+  select(Id, SalePrice)
+
+# Export
+write_csv(results, "Submissions/Custom_Model.csv")
 
@@
 
-Best custom model using caret
+Best custom model starting with the forward model and then adding in interactions
 ```{r}
-customModel = lm(SalePrice ~ I(MSSubClass^2) + log(LotArea) + Street + LotShape + LandContour + LotConfig + LandSlope + Neighborhood + Condition1 + Condition2 + BldgType + HouseStyle + OverallQual + OverallCond + YearBuilt + YearRemodAdd + RoofStyle + RoofMatl + ExterQual + ExterCond + Foundation + Heating + HeatingQC + CentralAir + `1stFlrSF` + `2ndFlrSF` + LowQualFinSF + GrLivArea + FullBath + HalfBath + BedroomAbvGr + KitchenAbvGr + TotRmsAbvGrd + Fireplaces + PavedDrive + WoodDeckSF + OpenPorchSF + EnclosedPorch + `3SsnPorch` + ScreenPorch + PoolArea + log(MiscVal) + MoSold + YrSold + SaleCondition, data = housesTrain)
-summary(fit)
+
+# housesTrainSub = housesTrain |>
+#   select(OverallQual, GrLivArea, Neighborhood, OverallCond, HouseStyle, YearBuilt, LotArea, RoofMatl, KitchenAbvGr, SaleCondition, Condition2, Foundation, Fireplaces, Heating, ExterQual, Condition1, PoolArea, ScreenPorch, WoodDeckSF, HeatingQC, CentralAir, BedroomAbvGr, FirstFlrSF, `2ndFlrSF`, Street, LandSlope, HalfBath, EnclosedPorch, SecondFlrSF, MiscVal, PavedDrive, `1stFlrSF`, BldgType, ExterCond, YearRemodAdd, SalePrice)
+
+housesTrainSub = housesTrain |>
+  select(OverallQual, GrLivArea, Neighborhood, OverallCond, HouseStyle, YearBuilt, LotArea, SalePrice)
+
+forwardResultsInteractions = ols_step_forward_p(lm(SalePrice ~ . + (.)^2, data = housesTrainSub), details = TRUE, progress = TRUE)
+
+plot(forwardResultsInteractions)
+
+forwardResultsInteractionsModel = forwardResultsInteractions$model
+summary(forwardResultsInteractionsModel)
+
+results = housesTest
+results$SalePrice = exp(predict(forwardResultsInteractionsModel, newdata = housesTest))
+results = results |>
+  select(Id, SalePrice)
+
+# Export
+write_csv(results, "Submissions/Custom_Model.csv")
+
+```
+
+Best custom model starting with the forward model, which was found to be the best between forward / backward / stepwise
+```{r}
+customModel = lm(SalePrice ~ OverallQual + GrLivArea + Neighborhood + OverallCond + HouseStyle + YearBuilt + LotArea + RoofMatl + KitchenAbvGr + SaleCondition + Condition2 + Foundation + Fireplaces + Heating + ExterQual + Condition1 + PoolArea + ScreenPorch + WoodDeckSF + HeatingQC + CentralAir + BedroomAbvGr + FirstFlrSF + `2ndFlrSF` + Street + LandSlope + HalfBath + EnclosedPorch + SecondFlrSF + MiscVal + PavedDrive + `1stFlrSF` + BldgType + ExterCond + YearRemodAdd, data = housesTrain)
+summary(customModel)
+
+#control = trainControl(method = "LOOCV", number = 10)
+control = trainControl(method = "cv")
+customModel2 = train(SalePrice ~ OverallQual + GrLivArea + Neighborhood + OverallCond + HouseStyle + YearBuilt + LotArea + RoofMatl + KitchenAbvGr + SaleCondition + Condition2 + Foundation + Fireplaces + Heating + ExterQual + Condition1 + PoolArea + ScreenPorch + WoodDeckSF + HeatingQC + CentralAir + BedroomAbvGr + FirstFlrSF + `2ndFlrSF` + Street + LandSlope + HalfBath + EnclosedPorch + SecondFlrSF + MiscVal + PavedDrive + `1stFlrSF` + BldgType + ExterCond + YearRemodAdd, data = housesTrain, method = "lm", trControl = control)
+
+summary(customModel2)
+
+results = housesTest
+results$SalePrice = exp(predict(customModel2, newdata = housesTest))
+results = results |>
+  select(Id, SalePrice)
+
+# Export
+write_csv(results, "Submissions/Custom_Model.csv")
+
+```
+
+Best custom model using cross validation
+```{r}
+# customModel = lm(SalePrice ~ I(MSSubClass^2) + log(LotArea) + Street + LotShape + LandContour + LotConfig + LandSlope + Neighborhood + Condition1 + Condition2 + BldgType + HouseStyle + OverallQual + OverallCond + YearBuilt + YearRemodAdd + RoofStyle + RoofMatl + ExterQual + ExterCond + Foundation + Heating + HeatingQC + CentralAir + `1stFlrSF` + `2ndFlrSF` + LowQualFinSF + GrLivArea + FullBath + HalfBath + BedroomAbvGr + KitchenAbvGr + TotRmsAbvGrd + Fireplaces + PavedDrive + WoodDeckSF + OpenPorchSF + EnclosedPorch + `3SsnPorch` + ScreenPorch + PoolArea + log(MiscVal) + MoSold + YrSold + SaleCondition, data = housesTrain)
+# summary(fit)
+
+customModel = lm(SalePrice ~ MSSubClass + LotArea + Street + LotShape + LandContour + LotConfig + LandSlope + Neighborhood + Condition1 + Condition2 + BldgType + HouseStyle + OverallQual + OverallCond + YearBuilt + YearRemodAdd + RoofStyle + RoofMatl + ExterQual + ExterCond + Foundation + Heating + HeatingQC + CentralAir + `1stFlrSF` + `2ndFlrSF` + LowQualFinSF + GrLivArea + FullBath + HalfBath + BedroomAbvGr + KitchenAbvGr + TotRmsAbvGrd + Fireplaces + PavedDrive + WoodDeckSF + OpenPorchSF + EnclosedPorch + `3SsnPorch` + ScreenPorch + PoolArea + MiscVal + MoSold + YrSold + SaleCondition, data = housesTrain)
+summary(customModel)
+
+customModel = lm(SalePrice ~ . +(.)^2, data = housesTrain)
+summary(customModel)
 
@@
 write_csv(results, "Submissions/Custom_Model.csv")
-
 ```
 
-Best custom model using external cross validation
+Custom model using MASS
 ```{r}
-customModel = lm(SalePrice ~ I(MSSubClass^2) + log(LotArea) + Street + LotShape + LandContour + LotConfig + LandSlope + Neighborhood + Condition1 + Condition2 + BldgType + HouseStyle + OverallQual + OverallCond + YearBuilt + YearRemodAdd + RoofStyle + RoofMatl + ExterQual + ExterCond + Foundation + Heating + HeatingQC + CentralAir + `1stFlrSF` + `2ndFlrSF` + LowQualFinSF + GrLivArea + FullBath + HalfBath + BedroomAbvGr + KitchenAbvGr + TotRmsAbvGrd + Fireplaces + PavedDrive + WoodDeckSF + OpenPorchSF + EnclosedPorch + `3SsnPorch` + ScreenPorch + PoolArea + log(MiscVal) + MoSold + YrSold + SaleCondition, data = housesTrain)
-summary(fit)
-
-#control = trainControl(method = "LOOCV", number = 10)
-control = trainControl(method = "cv")
-customModel = train(SalePrice ~ ., data = housesTrain, method = "lm", trControl = control)
-summary(customModel)
+initial = lm(SalePrice ~ ., data = housesTrain)
+stepModel = stepAIC(initial, direction = "both", trace = TRUE)
+summary(stepModel)
 
 results = housesTest
-results$SalePrice = exp(predict(customModel, newdata = housesTest))
+results$SalePrice = exp(predict(stepModel, newdata = housesTest))
 results = results |>
-  select(Id, SalePrice)
+  dplyr::select(Id, SalePrice)
 
@@
 write_csv(results, "Submissions/Custom_Model.csv")
-
 ```

### 2022-12-06 Analysis 2.qmd
Additional write up complete

@@
 
-# Retrieved from the output
-# forwardModel = lm(SalePrice ~ OverallQual + GrLivArea + Neighborhood + OverallCond + HouseStyle + YearBuilt + LotArea + RoofMatl + KitchenAbvGr + SaleCondition + Condition2 + Foundation + Fireplaces + Heating + ExterQual + Condition1 + PoolArea + ScreenPorch + WoodDeckSF + HeatingQC + CentralAir + BedroomAbvGr + FirstFlrSF + `2ndFlrSF` + Street + LandSlope + HalfBath + EnclosedPorch + SecondFlrSF + MiscVal + PavedDrive + `1stFlrSF` + BldgType + ExterCond + YearRemodAdd, data = housesTrain)
-
 plot(forwardResults)
@@
 backwardModel = backwardResults$model
+summary(backwardModel)
 
@@
 
-summary(backwardModel)
+plot(backwardModel)
 
@@
 
+plot(stepModel)
+
 results = housesTest
@@
 # Export
-write_csv(results, "Submissions/Both_Model.csv")
+write_csv(results, "Submissions/Step_Model.csv")
 ```
 
-# For the custom model use the variables determined from the forward selection and then try and improve the model using CV
+# Custom model using all variables - note, this was unable to achieve higher accuracy than the forward selection results
 ```{r}
+customModel = lm(SalePrice ~ I(MSSubClass^2) + I(LotArea)^2 + Street + LotShape + LandContour + LotConfig + LandSlope + Neighborhood + Condition1 + Condition2 + BldgType + HouseStyle + I(OverallQual^2) + OverallCond + I(YearBuilt^2) + YearRemodAdd + RoofStyle + RoofMatl + ExterQual + ExterCond + Foundation + Heating + HeatingQC + CentralAir + log(`1stFlrSF`) + `2ndFlrSF` + LowQualFinSF + I(GrLivArea^2) + FullBath + HalfBath + BedroomAbvGr + KitchenAbvGr + I(TotRmsAbvGrd^2) + I(Fireplaces^2) + PavedDrive + WoodDeckSF + OpenPorchSF + EnclosedPorch + `3SsnPorch` + ScreenPorch + PoolArea + MiscVal + MoSold + YrSold + SaleCondition, data = housesTrain)
 
-housesTrainSub = housesTrain |>
-  select(OverallQual, GrLivArea, Neighborhood, OverallCond, HouseStyle, YearBuilt, LotArea, RoofMatl, KitchenAbvGr, SaleCondition, Condition2, Foundation, Fireplaces, Heating, ExterQual, Condition1, PoolArea, ScreenPorch, WoodDeckSF, HeatingQC, CentralAir, BedroomAbvGr, FirstFlrSF, `2ndFlrSF`, Street, LandSlope, HalfBath, EnclosedPorch, SecondFlrSF, MiscVal, PavedDrive, `1stFlrSF`, BldgType, ExterCond, YearRemodAdd, SalePrice)
-
-# Define the number of subsets (or "folds") to use
-#train_control = trainControl(method = "cv", number = 5)
-train_control = trainControl(method = "LOOCV")
-
-#train the model
-customModel = train(SalePrice ~ ., data = housesTrainSub, method = "lm", trControl = train_control)
-print(customModel)
-summary(customModel)
-
-results = housesTest
-results$SalePrice = exp(predict(customModel, newdata = housesTest))
-results = results |>
-  select(Id, SalePrice)
-
-# Export
-write_csv(results, "Submissions/Custom_Model.csv")
-
-```
-
-Best custom model starting with the forward model and then adding in interactions
-```{r}
-
-# housesTrainSub = housesTrain |>
-#   select(OverallQual, GrLivArea, Neighborhood, OverallCond, HouseStyle, YearBuilt, LotArea, RoofMatl, KitchenAbvGr, SaleCondition, Condition2, Foundation, Fireplaces, Heating, ExterQual, Condition1, PoolArea, ScreenPorch, WoodDeckSF, HeatingQC, CentralAir, BedroomAbvGr, FirstFlrSF, `2ndFlrSF`, Street, LandSlope, HalfBath, EnclosedPorch, SecondFlrSF, MiscVal, PavedDrive, `1stFlrSF`, BldgType, ExterCond, YearRemodAdd, SalePrice)
-
-housesTrainSub = housesTrain |>
-  select(OverallQual, GrLivArea, Neighborhood, OverallCond, HouseStyle, YearBuilt, LotArea, SalePrice)
-
-forwardResultsInteractions = ols_step_forward_p(lm(SalePrice ~ . + (.)^2, data = housesTrainSub), details = TRUE, progress = TRUE)
-
-plot(forwardResultsInteractions)
-
-forwardResultsInteractionsModel = forwardResultsInteractions$model
-summary(forwardResultsInteractionsModel)
-
-results = housesTest
-results$SalePrice = exp(predict(forwardResultsInteractionsModel, newdata = housesTest))
-results = results |>
-  select(Id, SalePrice)
-
-# Export
-write_csv(results, "Submissions/Custom_Model.csv")
-
-```
-
-Best custom model starting with the forward model, which was found to be the best between forward / backward / stepwise
-```{r}
-customModel = lm(SalePrice ~ OverallQual + GrLivArea + Neighborhood + OverallCond + HouseStyle + YearBuilt + LotArea + RoofMatl + KitchenAbvGr + SaleCondition + Condition2 + Foundation + Fireplaces + Heating + ExterQual + Condition1 + PoolArea + ScreenPorch + WoodDeckSF + HeatingQC + CentralAir + BedroomAbvGr + FirstFlrSF + `2ndFlrSF` + Street + LandSlope + HalfBath + EnclosedPorch + SecondFlrSF + MiscVal + PavedDrive + `1stFlrSF` + BldgType + ExterCond + YearRemodAdd, data = housesTrain)
-summary(customModel)
-
-#control = trainControl(method = "LOOCV", number = 10)
-control = trainControl(method = "cv")
-customModel2 = train(SalePrice ~ OverallQual + GrLivArea + Neighborhood + OverallCond + HouseStyle + YearBuilt + LotArea + RoofMatl + KitchenAbvGr + SaleCondition + Condition2 + Foundation + Fireplaces + Heating + ExterQual + Condition1 + PoolArea + ScreenPorch + WoodDeckSF + HeatingQC + CentralAir + BedroomAbvGr + FirstFlrSF + `2ndFlrSF` + Street + LandSlope + HalfBath + EnclosedPorch + SecondFlrSF + MiscVal + PavedDrive + `1stFlrSF` + BldgType + ExterCond + YearRemodAdd, data = housesTrain, method = "lm", trControl = control)
-
-summary(customModel2)
-
-results = housesTest
-results$SalePrice = exp(predict(customModel2, newdata = housesTest))
-results = results |>
-  select(Id, SalePrice)
-
-# Export
-write_csv(results, "Submissions/Custom_Model.csv")
-
-```
-
-Best custom model using cross validation
-```{r}
-# customModel = lm(SalePrice ~ I(MSSubClass^2) + log(LotArea) + Street + LotShape + LandContour + LotConfig + LandSlope + Neighborhood + Condition1 + Condition2 + BldgType + HouseStyle + OverallQual + OverallCond + YearBuilt + YearRemodAdd + RoofStyle + RoofMatl + ExterQual + ExterCond + Foundation + Heating + HeatingQC + CentralAir + `1stFlrSF` + `2ndFlrSF` + LowQualFinSF + GrLivArea + FullBath + HalfBath + BedroomAbvGr + KitchenAbvGr + TotRmsAbvGrd + Fireplaces + PavedDrive + WoodDeckSF + OpenPorchSF + EnclosedPorch + `3SsnPorch` + ScreenPorch + PoolArea + log(MiscVal) + MoSold + YrSold + SaleCondition, data = housesTrain)
-# summary(fit)
-
-customModel = lm(SalePrice ~ MSSubClass + LotArea + Street + LotShape + LandContour + LotConfig + LandSlope + Neighborhood + Condition1 + Condition2 + BldgType + HouseStyle + OverallQual + OverallCond + YearBuilt + YearRemodAdd + RoofStyle + RoofMatl + ExterQual + ExterCond + Foundation + Heating + HeatingQC + CentralAir + `1stFlrSF` + `2ndFlrSF` + LowQualFinSF + GrLivArea + FullBath + HalfBath + BedroomAbvGr + KitchenAbvGr + TotRmsAbvGrd + Fireplaces + PavedDrive + WoodDeckSF + OpenPorchSF + EnclosedPorch + `3SsnPorch` + ScreenPorch + PoolArea + MiscVal + MoSold + YrSold + SaleCondition, data = housesTrain)
-summary(customModel)
-
-customModel = lm(SalePrice ~ . +(.)^2, data = housesTrain)
-summary(customModel)
-
-#control = trainControl(method = "LOOCV", number = 10)
-control = trainControl(method = "cv")
-customModel = train(SalePrice ~ ., data = housesTrain, method = "lm", trControl = control)
 summary(customModel)
@@
 ```
-
-Custom model using MASS
-```{r}
-initial = lm(SalePrice ~ ., data = housesTrain)
-stepModel = stepAIC(initial, direction = "both", trace = TRUE)
-summary(stepModel)
-
-results = housesTest
-results$SalePrice = exp(predict(stepModel, newdata = housesTest))
-results = results |>
-  dplyr::select(Id, SalePrice)
-
-# Export
-write_csv(results, "Submissions/Custom_Model.csv")
-```

### 2022-12-06 MSDS 6371 Project - Yule and Kondakindi.qmd
Analysis 2: Moved code from main markdown file into separate one to focus on Analysis 2 question
Additional write up complete

@@
 
+This analysis focuses on applying multi-linear regression techniques to predict the sales price of homes in the Ames, Iowa area. Two specific questions of interest (QOI) were identified:
+
+1.
+Focusing on 3 specific neighborhoods in Ames, what is the best linear regression model that can be constructed solely utilizing the square footage of the living room and what can be interpreted from that model?
+2.  Utilizing all available variables and focusing on every neighborhood, what are the best linear regression models that can be constructed?
+
 # Data Description
+
+For this analysis, housing data was obtained between the years of 2006 and 2010 from 1,460 homes in the Ames area.
+Approximately 80 variables were collected for each home identifying various features that may contribute to a homes final sale price.
+
+Upon analysis of the data, numerous variables were found to have a high amount of associated missing values and were subsequently removed from the analysis.
+
+Many numerical variables in the available data showed high variance and were therefore given a log transformation to better perform in a linear regression model.
+The full list of variables that received a log transformation can be found below.
+It should be noted that sales price was among the variables.
+
+- LotArea
+- FirstFlrSF
+- SecondFlrSF
+- GrLivArea
+- WoodDeckSF
+- OpenPorchSF
+- EnclosedPorch
+- ThirdSsnPorch
+- ScreenPorch
+- PoolArea
+- MiscVal
+- SalePrice
 
@@
 
-Brief introduction to the questions of interest and the setting of the problem.
+(Brief introduction to the questions of interest and the setting of the problem.)
 
@@
 
-Residual Plots Influential point analysis (Cook's D and Leverage) Make sure to address each assumption.
+(Residual Plots Influential point analysis (Cook's D and Leverage) Make sure to address each assumption.)
 
@@
 
-Adj R2 Internal CV Press
+(Adj R2 Internal CV Press)
 
@@
 
-Estimates Interpretation Confidence Intervals
+(Estimates Interpretation Confidence Intervals)
 
@@
 
-A short summary of the analysis.
+(A short summary of the analysis.)
 
@@
 
+A web application was created to provide access to the model and can be found at the link below:
+
+[www.google.com](https://www.google.com)
+
 ### Question 2
 
-#### Restatement of Problem
+For the second question of interest, the objective was to utilize all available data and variables to create the best linear regression model possible for predicting home sales prices.
+There are numerous techniques available for creating multi-linear regression models.
+This QOI focused on applying forward selection, backwards elimination, and stepwise selection to produce the highest performing models.
+Additionally, a final custom model was fit using cross validation techniques.
 
@@
 
-Type of Selection Stepwise Forward Backward Custom
+The olsrr package in R was utilized to perform each of the 3 stepping regression fittings.
+Minimizing the Akaike Information Criterion (AIC) was used as the model criteria.
+The following linear forms below were identified to produce the lowest AIC for forward, backward, and step selection respectively.
+A custom model was also attempted, however, it was unable to achieve higher accuracy than the forward selection model and thus was discarded.
+
+Forward selection:
+
+SalePrice = OverallQual + GrLivArea + Neighborhood + OverallCond + HouseStyle + YearBuilt + LotArea + RoofMatl + KitchenAbvGr + SaleCondition + Condition2 + Foundation + Fireplaces + Heating + ExterQual + Condition1 + PoolArea + ScreenPorch + WoodDeckSF + HeatingQC + CentralAir + BedroomAbvGr + FirstFlrSF + `2ndFlrSF` + Street + LandSlope + HalfBath + EnclosedPorch + SecondFlrSF + MiscVal + PavedDrive + `1stFlrSF` + BldgType + ExterCond + YearRemodAdd
+
+Backwards elimination:
+
+SalePrice = LotArea + Street + LotConfig + LandSlope + Neighborhood + Condition1 + Condition2 + BldgType + OverallQual + OverallCond + YearBuilt + YearRemodAdd + RoofMatl + ExterQual + ExterCond + Foundation + Heating + HeatingQC + CentralAir + `1stFlrSF` + `2ndFlrSF` + GrLivArea + HalfBath + BedroomAbvGr + KitchenAbvGr + Fireplaces + PavedDrive + WoodDeckSF + EnclosedPorch + ScreenPorch + PoolArea + MiscVal + SaleCondition + FirstFlrSF + SecondFlrSF
+
+Stepwise selection:
+
+SalePrice = OverallQual + GrLivArea + Neighborhood + OverallCond + YearBuilt + LotArea + RoofMatl + KitchenAbvGr + SaleCondition + Condition2 + Foundation + Fireplaces + Heating + ExterQual + Condition1 + PoolArea + ScreenPorch + WoodDeckSF + HeatingQC + CentralAir + BedroomAbvGr + FirstFlrSF + `2ndFlrSF` + Street + LandSlope + HalfBath + EnclosedPorch + SecondFlrSF + MiscVal + PavedDrive + BldgType + `1stFlrSF` + ExterCond + LotConfig + YearRemodAdd
 
@@
 
-Residual Plots Influential point analysis (Cook's D and Leverage) Make sure to address each assumption
+(Residual Plots Influential point analysis (Cook's D and Leverage) Make sure to address each assumption)
+
+![](Model Metrics/Collage.png)
 
@@
 
-Adj R2 Internal CV Press\
-Kaggle Score
+The table below summarizes the key performance metrics associated with each of the three models produced. Each model was found to have the same Adjusted R Squared metric without rounding errors.
+
+|     Model                 |  Adjusted  R2   |   CV Press    |   Kaggle Score    |
+|---------------------------|-----------------|---------------|-------------------|
+|   Forward Selection       | 0.9094          |   X           |   0.13496         |
+|   Backward Elimination    | 0.9094          |   X           |   0.13554         |
+|   Stepwise Selection      | 0.9094          |   X           |   0.13554         |
 
@@
 
-A short summary of the analysis
+(A short summary of the analysis)
 
@@
 
-```{r}
-library(tidyverse)
-```
-
-Read training and test data and perform any cleaning steps
-```{r}
-housesTrain = read_csv("train.csv")
-housesTest = read_csv("test.csv")
-
-remove(houses)
-```
+The entire code for this analysis can be found below:

### 2022-12-08 Analysis 2.qmd
Writing updates

@@
 
+Define a function to determine the PRESS of each model
+```{r}
+PRESS = function(model) {
+    i = residuals(model)/(1 - lm.influence(model)$hat)
+    sum(i^2)
+}
+```
+
 Forward step regression using olss
 ```{r}
+# Remove high influence rows
+# housesTrain2 = housesTrain |>
+#   filter(!row_number() %in% c(121, 251, 272, 326, 584, 1004, 1231, 1276, 1299, 1322))
+
 forwardResults = ols_step_forward_aic(lm(SalePrice ~ ., data = housesTrain), details = TRUE, progress = TRUE)
@@
 
+# Plot the stepping results impact on AIC
 plot(forwardResults)
 
+# Plot the residual / leverage functions
 plot(forwardModel)
-#performance::check_model(forwardModel, check = c("linearity", "normality", "qq"))
 
+# Calculate the PRESS for the model
+tibble(resids = residuals(forwardModel), hats = tidy(hatvalues(forwardModel))$x) |>
+  filter(hats != 1) |>
+  mutate(i = (resids / (1-hats))^2) |>
+  select(i) |>
+  pull() |>
+  sum()
+
+# Calculate values against the test data set
 results = housesTest
@@
 
-# Export
+# Export and submit to Kaggle
 write_csv(results, "Submissions/Forward_Model.csv")
@@
 
-# backwardModel = lm(SalePrice ~ LotArea + Street + LotConfig + LandSlope + Neighborhood + Condition1 + Condition2 + BldgType + OverallQual + OverallCond + YearBuilt + YearRemodAdd + RoofMatl + ExterQual + ExterCond + Foundation + Heating + HeatingQC + CentralAir + `1stFlrSF` + `2ndFlrSF` + GrLivArea + HalfBath + BedroomAbvGr + KitchenAbvGr + Fireplaces + PavedDrive + WoodDeckSF + EnclosedPorch + ScreenPorch + PoolArea + MiscVal + SaleCondition + FirstFlrSF + SecondFlrSF, data = housesTrain)
-
+# Plot the stepping results impact on AIC
 plot(backwardResults)
 
+# Plot the residual / leverage functions
 plot(backwardModel)
 
+# Calculate the PRESS for the model
+tibble(resids = residuals(backwardModel), hats = tidy(hatvalues(backwardModel))$x) |>
+  filter(hats != 1) |>
+  mutate(i = (resids / (1-hats))^2) |>
+  select(i) |>
+  pull() |>
+  sum()
+
+# Calculate values against the test data set
 results = housesTest
@@
 
-# Export
+# Export and submit to Kaggle
 write_csv(results, "Submissions/Backward_Model.csv")
@@
 stepModel = lm(SalePrice ~ OverallQual + GrLivArea + Neighborhood + OverallCond + YearBuilt + LotArea + RoofMatl + KitchenAbvGr + SaleCondition + Condition2 + Foundation + Fireplaces + Heating + ExterQual + Condition1 + PoolArea + ScreenPorch + WoodDeckSF + HeatingQC + CentralAir + BedroomAbvGr + FirstFlrSF + `2ndFlrSF` + Street + LandSlope + HalfBath + EnclosedPorch + SecondFlrSF + MiscVal + PavedDrive + BldgType + `1stFlrSF` + ExterCond + LotConfig + YearRemodAdd, data = housesTrain)
-
+# Plot the stepping results impact on AIC
 plot(stepResults)
 
-summary(stepModel)
-
+# Plot the residual / leverage functions
 plot(stepModel)
 
+# Calculate the PRESS for the model
+tibble(resids = residuals(stepModel), hats = tidy(hatvalues(stepModel))$x) |>
+  filter(hats != 1) |>
+  mutate(i = (resids / (1-hats))^2) |>
+  select(i) |>
+  pull() |>
+  sum()
+
+# Calculate values against the test data set
 results = housesTest
@@
 
-# Export
+# Export and submit to Kaggle
 write_csv(results, "Submissions/Step_Model.csv")
@@
 
+# Calculate values against the test data setresults = housesTest
 results = housesTest
@@
 
-# Export
+# Export and submit to Kaggle
 write_csv(results, "Submissions/Custom_Model.csv")

### 2022-12-08 MSDS 6371 Project - Yule and Kondakindi.qmd
Minor updates
Writing updates

@@
 Focusing on 3 specific neighborhoods in Ames, what is the best linear regression model that can be constructed solely utilizing the square footage of the living room and what can be interpreted from that model?
-2.  Utilizing all available variables and focusing on every neighborhood, what are the best linear regression models that can be constructed?
+2.
+Utilizing all available variables and focusing on every neighborhood, what are the best linear regression models that can be constructed?
+For this QOI, submissions will be judged based on a [Kaggle competition](https://www.kaggle.com/competitions/house-prices-advanced-regression-techniques), using produced model predictions to achieve the lowest root mean squared error.
 
@@
 
-For this analysis, housing data was obtained between the years of 2006 and 2010 from 1,460 homes in the Ames area.
+For this analysis, housing data from the Ames area was supplied between the years of 2006 and 2010 from 1,460 homes.
 Approximately 80 variables were collected for each home identifying various features that may contribute to a homes final sale price.
+The variables ranged in nature from quantitative values like square footage of certain parts of the home, to qualitative values like quality rankings and building materials.
 
-Upon analysis of the data, numerous variables were found to have a high amount of associated missing values and were subsequently removed from the analysis.
+Upon analysis of the data, numerous variables were found to have a high number of missing values and were subsequently removed from the analysis.
 
-Many numerical variables in the available data showed high variance and were therefore given a log transformation to better perform in a linear regression model.
+Many numerical variables in the available data demonstrated high variance.
+Log transformations were used to reduce the variance and ultimately improve the performance of linear regression models.
 The full list of variables that received a log transformation can be found below.
-It should be noted that sales price was among the variables.
+It should be noted that sales price (the variable of interest) was among the variables recieving a log transformation.
 
@@
 
-(Brief introduction to the questions of interest and the setting of the problem.)
-
 #### Restatement of the Problem
@@
 
+For the first question of interest, the objective was to utilize only the greater living room square footage of a house to help predict the sales price in the three Ames neighborhoods of Brookside, Edwards, and North Ames.
+
+Limiting the initial dataset to only these 3 neighborhoods results in 383 observations with 58, 100, and 225 for Brookside, Edwards, and North Ames respectively.
+
 #### Build and Fit the Model
+
+Three models were fit to the data:
+
+1. Average per neighborhood - A model which assumes only the average sales price for each neighborhood.
+2. Varying y-intercept / same slopes - A model which allows for varying y-intercepts across each neighborhood while maintaining the same relationship in sales price and living room square footage.
+3. Varying y-intercept / varying slopes - A model which allows for varying y-intercepts across each neighborhood in addition to varying relationships in sales price and living room square footage.
 
@@
 
-(Adj R2 Internal CV Press)
+The table below summarizes the key performance metrics associated with each of the three models produced.
+Overall the third model, which allows for varying y-intercepts and slopes across each neighborhood, outperforms the other 2 models.
+
+|     Model                               |  Adjusted  R2   |   CV PRESS    |
+|-----------------------------------------|-----------------|---------------|
+|   Average per neighborhood              | 0.0880          |   26.44       |
+|   Varying y-intercept / same slope      | 0.4857          |   15.00       |
+|   Varying y-intercepts / varying slopes | 0.5056          |   14.61       |
 
@@
 This QOI focused on applying forward selection, backwards elimination, and stepwise selection to produce the highest performing models.
-Additionally, a final custom model was fit using cross validation techniques.
 
@@
 
-The table below summarizes the key performance metrics associated with each of the three models produced. Each model was found to have the same Adjusted R Squared metric without rounding errors.
+The table below summarizes the key performance metrics associated with each of the three models produced.
+Each model was found to have the same Adjusted R Squared metric within rounding errors.
+According to the predictive residual sum of squares metric, the stepwise selection model performed the best.
+When used to predict unknown home sales prices as part of the Kaggle competition, the Forward Selection model was found to achieve the best (lowest) Kaggle score of 0.13496.
+A custom model was excluded from this table as it was unable to achieve higher performance than the Forward and Stepwise selection methods.
 
-|     Model                 |  Adjusted  R2   |   CV Press    |   Kaggle Score    |
+|     Model                 |  Adjusted  R2   |   CV PRESS    |   Kaggle Score    |
 |---------------------------|-----------------|---------------|-------------------|
-|   Forward Selection       | 0.9094          |   X           |   0.13496         |
-|   Backward Elimination    | 0.9094          |   X           |   0.13554         |
-|   Stepwise Selection      | 0.9094          |   X           |   0.13554         |
+|   Forward Selection       | 0.9094          |   26.12       |   0.13496         |
+|   Backward Elimination    | 0.9094          |   25.95       |   0.13554         |
+|   Stepwise Selection      | 0.9094          |   25.94       |   0.13554         |
 
@@
 
-(A short summary of the analysis)
+In this analysis, a data set of home sales prices from Ames, Iowa was explored using multi-linear regression techniques.
+A model was produced to predict the sales price of homes specifically in the North Ames, Edwards, and Brk Side neighborhoods using only the greater living area square footage.
+That model achieved an RSquared metric of XX, meaning XX% of the variance in home prices in those 3 neighborhoods can be explained soley by the living room squared footage.
+
+Finally, 4 additional models were produced to predict homes sales prices based on all variables available.
+Various multi-linear regression stepping approaches were used and the forward selection method was found to perform the best overall with an adjusted RSquared metric of 0.9094 meaning over 90% of the variance in home prices could be explained by the variables and associated model.
+Additionally, a Kaggle competition score of 0.13496 was achieved by the model on a blind test.
 

### 2022-12-12 Analysis 1.qmd
Updates

---
title: "Analysis 1"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console
---

Define a function to determine the PRESS of each model
```{r}
PRESS = function(model) {
    i = residuals(model)/(1 - lm.influence(model)$hat)
    sum(i^2)
}
```

Create models for Sales Price for the neighborhood Brookside, Edwards, and North Ames neighborhoods.
```{r}

housesTrain |>
  ggplot(aes(x = GrLivArea, y = SalePrice)) +
  geom_point()

housesTrainNeighborhood = housesTrain |>
  filter(Neighborhood %in% c("BrkSide", "Edwards", "NAmes"))

housesTrainNeighborhood$Neighborhood = factor(housesTrainNeighborhood$Neighborhood, levels = c("NAmes", "BrkSide", "Edwards"))

housesTrainNeighborhood |>
  group_by(Neighborhood) |>
  summarize(n = n())

housesTrainNeighborhood |>
  ggplot(aes(x = GrLivArea, y = SalePrice, color = Neighborhood)) +
  geom_point() +
  facet_wrap(~Neighborhood) +
  theme(legend.position="none") +
  xlab("Log Value of Living Area Square Footage") +
  ylab("Log Value of Sales Price") +
  ggtitle("Home Sales Price and Living Area")

model1 = lm(SalePrice ~ Neighborhood, data = housesTrainNeighborhood)
summary(model1)
PRESS(model1)

model2 = lm(SalePrice ~ Neighborhood + GrLivArea, data = housesTrainNeighborhood)
summary(model2)
PRESS(model2)

model3 = lm(SalePrice ~ Neighborhood + GrLivArea + Neighborhood:GrLivArea, data = housesTrainNeighborhood)
summary(model3)
PRESS(model3)

plot(model3)

housesTrainNeighborhood |>
  select(Neighborhood, GrLivArea, SalePrice) |>
  mutate(SalePricePredicted = predict(model3, newdata = tibble(Neighborhood = Neighborhood, GrLivArea = GrLivArea))) |>
  ggplot() +
  geom_point(aes(x = exp(GrLivArea), y = exp(SalePrice), color = Neighborhood)) +
  geom_line(aes(x = exp(GrLivArea), y = exp(SalePricePredicted), color = Neighborhood)) +
  facet_wrap(~Neighborhood) +
  theme(legend.position="none")

```

### 2022-12-12 Analysis 2.qmd
Updates

@@
 Load required libraries
+
 ```{r}
@@
 Read training and test data, then drop any columns containing missing values
+
 ```{r}
@@
 
-The following variables can be log transformed in order to reduce their variance:
-- LotArea
-- FirstFlrSF
-- SecondFlrSF
-- GrLivArea
-- WoodDeckSF
-- OpenPorchSF
-- EnclosedPorch
-- ThirdSsnPorch
-- ScreenPorch
-- PoolArea
-- MiscVal
-- SalePrice
+The following variables can be log transformed in order to reduce their variance: - LotArea - FirstFlrSF - SecondFlrSF - GrLivArea - WoodDeckSF - OpenPorchSF - EnclosedPorch - ThirdSsnPorch - ScreenPorch - PoolArea - MiscVal - SalePrice
 
 Additionally, MoSold variable was identified as needing to be a factor instead of an integer
+
 ```{r}
@@
 Define a function to determine the PRESS of each model
+
 ```{r}
@@
 Forward step regression using olss
+
 ```{r}
@@
 Backward step regression
+
 ```{r}
@@
 Stepwise regression
+
 ```{r}
@@
 # Custom model using all variables - note, this was unable to achieve higher accuracy than the forward selection results
+
 ```{r}

### 2022-12-12 Shinny App/AmesSalesPrices/server.R

@@ -1,0 +1,90 @@
+#
+# This is the server logic of a Shiny web application. You can run the
+# application by clicking 'Run App' above.
+#
+# Find out more about building applications with Shiny here:
+#
+#    http://shiny.rstudio.com/
+#
+
+library(shiny)
+library(tidyverse)
+
+housesTrain = tibble(
+  Neighborhood = c("BrkSide", "NAmes", "BrkSide", "NAmes", "NAmes", "NAmes", "NAmes", 
+                   "BrkSide", "NAmes", "NAmes", "NAmes", "Edwards", "NAmes", "NAmes", 
+                   "BrkSide", "NAmes", "NAmes", "NAmes", "NAmes", "NAmes", "NAmes", 
+                   "BrkSide", "NAmes", "NAmes", "NAmes", "NAmes", "Edwards", "Edwards", 
+                   "NAmes", "Edwards", "Edwards", "NAmes", "NAmes", "NAmes", "NAmes", 
+                   "NAmes", "NAmes", "NAmes", "NAmes", "Edwards", "BrkSide", "BrkSide", 
+                   "Edwards", "NAmes", "Edwards", "NAmes", "NAmes", "Edwards", 
+                   "Edwards", "BrkSide", "NAmes", "Edwards", "NAmes", "Edwards", 
+                   "NAmes", "NAmes", "Edwards", "Edwards", "NAmes", "NAmes", "NAmes", 
+                   "Edwards", "NAmes", "BrkSide", "NAmes", "NAmes", "NAmes", "Edwards", 
+                   "NAmes", "BrkSide", "NAmes", "NAmes", "BrkSide", "Edwards", "NAmes", 
+                   "Edwards", "NAmes", "NAmes", "NAmes", "BrkSide", "NAmes", "NAmes", 
+                   "NAmes", "NAmes", "BrkSide", "NAmes", "NAmes", "Edwards", "BrkSide", 
+                   "Edwards", "NAmes", "NAmes", "NAmes", "NAmes", "NAmes", "Edwards", 
+                   "Edwards", "Edwards", "Edwards", "NAmes", "BrkSide", "Edwards", 
+                   "NAmes", "Edwards", "Edwards", "NAmes", "NAmes", "NAmes", "BrkSide", 
+                   "Edwards", "Edwards", "BrkSide", "Edwards", "NAmes", "BrkSide", 
+                   "NAmes", "NAmes", "Edwards", "NAmes", "NAmes", "NAmes", "NAmes", 
[…]

### 2022-12-12 Shinny App/AmesSalesPrices/ui.R

@@ -1,0 +1,33 @@
+#
+# This is the user-interface definition of a Shiny web application. You can
+# run the application by clicking 'Run App' above.
+#
+# Find out more about building applications with Shiny here:
+#
+#    http://shiny.rstudio.com/
+#
+
+library(shiny)
+
+# Define UI for application that draws a histogram
+shinyUI(fluidPage(
+
+    # Application title
+    titlePanel("Ames Housing Data"),
+    
+    # Sidebar with a slider input for number of bins
+    sidebarLayout(
+        sidebarPanel(
+          radioButtons(
+            "neighborhood",
+            "Neighborhood",
+            choices = c("NAmes", "BrkSide", "Edwards")
+          )
+        ),
+
+        # Show a plot of the generated distribution
+        mainPanel(
+            plotOutput("plot")
+        )
+    )
+))

### 2022-12-12 MSDS 6371 Project - Yule and Kondakindi.qmd
Updates

@@
 
-This analysis focuses on applying multi-linear regression techniques to predict the sales price of homes in the Ames, Iowa area. Two specific questions of interest (QOI) were identified:
+This analysis focuses on applying multi-linear regression techniques to predict the sales price of homes in the Ames area of Iowa. Two specific questions of interest (QOI) were identified:
 
-1.
-Focusing on 3 specific neighborhoods in Ames, what is the best linear regression model that can be constructed solely utilizing the square footage of the living room and what can be interpreted from that model?
+1.  Focusing on 3 specific neighborhoods in Ames, what is the relationship between a home's square footage and it's sales price.
 2.
@@
 For this analysis, housing data from the Ames area was supplied between the years of 2006 and 2010 from 1,460 homes.
-Approximately 80 variables were collected for each home identifying various features that may contribute to a homes final sale price.
+Approximately 80 variables were collected for each home identifying various features that may contribute to a home's final sale price.
 The variables ranged in nature from quantitative values like square footage of certain parts of the home, to qualitative values like quality rankings and building materials.
@@
 Many numerical variables in the available data demonstrated high variance.
-Log transformations were used to reduce the variance and ultimately improve the performance of linear regression models.
+Log transformations were used to reduce the variance and ultimately improve the performance of any linear regression models.
 The full list of variables that received a log transformation can be found below.
-It should be noted that sales price (the variable of interest) was among the variables recieving a log transformation.
+It should be noted that sales price (the variable of interest) was among the variables receiving a log transformation.
 
@@
 
-(Where did the data come from? How big is it? How many observations? Where can we find out more? What are the specific variables that we need to know with respect to your analysis?)
+For the first question of interest, the objective was to utilize only a home's square footage to help predict the sales price in the three neighborhoods of North Ames, Brookside, and Edwards.
 
-For the first question of interest, the objective was to utilize only the greater living room square footage of a house to help predict the sales price in the three Ames neighborhoods of Brookside, Edwards, and North Ames.
+Limiting the initial dataset to only these 3 neighborhoods results in 383 observations with 225, 58, and 100 for North Ames, Brookside, and Edwards respectively.
 
-Limiting the initial dataset to only these 3 neighborhoods results in 383 observations with 58, 100, and 225 for Brookside, Edwards, and North Ames respectively.
+Utilizing the log transformations performed in the section above, helps to improve the linear relationships between square footage and sales price for the 3 neighborhoods.
+
+![](Sales_Price_and_Living_Area.png)
 
@@
 
-1. Average per neighborhood - A model which assumes only the average sales price for each neighborhood.
+1. Average per neighborhood - A model which fits the average sales price for each neighborhood.
 2. Varying y-intercept / same slopes - A model which allows for varying y-intercepts across each neighborhood while maintaining the same relationship in sales price and living room square footage.
 3. Varying y-intercept / varying slopes - A model which allows for varying y-intercepts across each neighborhood in addition to varying relationships in sales price and living room square footage.
-
-#### Checking the Assumptions
-
-(Residual Plots Influential point analysis (Cook's D and Leverage) Make sure to address each assumption.)
 
@@
 
+#### Checking the Assumptions
+
+The figure below summarizes the residual and influential point analysis conducted on the final model.
+Overall, the model appears to meet the basic assumptions required of linear regression as indicated by a random scattering of the residuals.
+Several high leverage points were found in the influential point analysis, however, no points were removed as they could not confidently be assumed to be erroneous data.
+
+![](Model Metrics/Collage_Analysis1.png)
+
 #### Parameters
 
-(Estimates Interpretation Confidence Intervals)
+The table below summarizes the determined regression coefficients and associated confidence intervals. The reference for y-intercept and slope values is the North Ames neighborhood.
 
-#### Conclusion
+Based on these results, it can be determined that homes in the Brookside and Edwards neighborhoods have lower starting values for homes with low square footage, however, homes in those neighborhoods tend to increase in sales price values more sharply as square footage increases.
+More specifically, a doubling of square footage in each neighborhood is associated with a multiplicative increase in sales price of 39%, 76%, and 43% for North Ames, Brookside, and Edwards respectively.
 
-(A short summary of the analysis.)
+![](Model Metrics/Analysis1.png)
 
@@
 
-A web application was created to provide access to the model and can be found at the link below:
+To further explore the relationships between home square footage and sales price in these neighborhoods, a web application was created which can be found at the link below.
+Also, a link to this report can be found at the link below.
 
-[www.google.com](https://www.google.com)
+[Shinny App](https://ayule89.shinyapps.io/AmesSalesPrices/)
+
+[Link to Andrew's Personal Website](https://ayule89.github.io/MSDS_6371_Project_Yule_and_Kondakindi.pdf)
+
+[Link to Krithika's Personal Website](https://krithkonda.github.io/MSDS_6371_Project_Yule_and_Kondakindi.pdf)
 
@@
 The following linear forms below were identified to produce the lowest AIC for forward, backward, and step selection respectively.
-A custom model was also attempted, however, it was unable to achieve higher accuracy than the forward selection model and thus was discarded.
+A custom model was also attempted, however, it was unable to achieve higher accuracy than the preceding 3 models and thus was discarded.
 
@@
 
-(Residual Plots Influential point analysis (Cook's D and Leverage) Make sure to address each assumption)
+The figure below summarizes the residual and influential point analysis conducted on each of the 3 models.
+Overall, each model appears to meet the assumptions required of linear regression as indicated by a random scattering of the residuals.
+Several high leverage points were found in the influential point analysis, however, no points were removed as they could not confidently be assumed to be erroneous data.
 
@@
 In this analysis, a data set of home sales prices from Ames, Iowa was explored using multi-linear regression techniques.
-A model was produced to predict the sales price of homes specifically in the North Ames, Edwards, and Brk Side neighborhoods using only the greater living area square footage.
-That model achieved an RSquared metric of XX, meaning XX% of the variance in home prices in those 3 neighborhoods can be explained soley by the living room squared footage.
+A model was produced to predict the sales price of homes specifically in the North Ames, Brookside, and Edwards neighborhoods using only the square footage.
+That model achieved an RSquared metric of 0.51, meaning 51% of the variance in home prices in those 3 neighborhoods can be explained by the square footage.
 
-Finally, 4 additional models were produced to predict homes sales prices based on all variables available.
-Various multi-linear regression stepping approaches were used and the forward selection method was found to perform the best overall with an adjusted RSquared metric of 0.9094 meaning over 90% of the variance in home prices could be explained by the variables and associated model.
+Finally, 3 additional models were produced to predict homes sales prices based on all variables available.
+Various multi-linear regression stepping approaches were used and the forward selection method was found to perform the best overall with an adjusted RSquared metric of 0.91 meaning 91% of the variance in home prices could be explained by the variables and associated model.
 Additionally, a Kaggle competition score of 0.13496 was achieved by the model on a blind test.
@@
 The entire code for this analysis can be found below:
+
+Load required libraries
+
+```{r}
+#| eval: false
+#| output: false
+
+library(MASS)
+library(olsrr)
+library(caret)
+library(terra)
+library(performance)
+library(tidyverse)
+library(broom)
+```
+
+Read training and test data, then drop any columns containing missing values
+
+```{r}
+#| eval: false
+#| output: false
+
+housesTrain = read_csv("train.csv")
+housesTest = read_csv("test.csv")
+
+# Remove ID column only from the training set
+housesTrain = dplyr::select(housesTrain, -Id)
+
+# Find which columns have missing values between the training and test data sets
+colsToDrop = housesTrain |>
+  select_if(function(x) any(is.na(x))) |>
+  colnames()
+colsToDrop2 = housesTest |>
+  select_if(function(x) any(is.na(x))) |>
+  colnames()
+
+# Remove any of the columns that had missing values
+housesTrain =  housesTrain |>
+  dplyr::select(-starts_with(c(colsToDrop, colsToDrop2)))
+housesTest =  housesTest |>
+  dplyr::select(-starts_with(c(colsToDrop, colsToDrop2)))
+```
+
+```{r}
+#| eval: false
+#| output: false
+
+# Screen for numerical variables that should be log-transformed
+housesTrain |> 
+  dplyr::select(where(is.numeric)) |>
+  gather() |>
+  ggplot(aes(x = value)) + 
+    facet_wrap(~key, scales = "free") + 
+    geom_histogram()
+```
+
+The following variables can be log transformed in order to reduce their variance: - LotArea - FirstFlrSF - SecondFlrSF - GrLivArea - WoodDeckSF - OpenPorchSF - EnclosedPorch - ThirdSsnPorch - ScreenPorch - PoolArea - MiscVal - SalePrice
+
+Additionally, MoSold variable was identified as needing to be a factor instead of an integer
+
+```{r}
+#| eval: false
+#| output: false
+
+# MoSold should be a factor instead of integer
+housesTrain$MoSold = factor(housesTrain$MoSold, levels = seq(1, 12))
+housesTest$MoSold = factor(housesTest$MoSold, levels = seq(1, 12))
+
+# First ensure that any variables that need log transformations are set
+#to a minimum value of 1 instead of 0. Then, take the log of the values.
+housesTrain = housesTrain |>
+  mutate(
+    LotArea = log(clamp(LotArea, lower = 1)),
+    FirstFlrSF = log(clamp(`1stFlrSF`, lower = 1)),
+    SecondFlrSF = log(clamp(`2ndFlrSF`, lower = 1)),
+    GrLivArea = log(clamp(GrLivArea, lower = 1)),
+    WoodDeckSF = log(clamp(WoodDeckSF, lower = 1)),
+    OpenPorchSF = log(clamp(OpenPorchSF, lower = 1)),
+    EnclosedPorch = log(clamp(EnclosedPorch, lower = 1)),
+    ThirdSsnPorch = log(clamp(`3SsnPorch`, lower = 1)),
+    ScreenPorch = log(clamp(ScreenPorch, lower = 1)),
+    PoolArea = log(clamp(PoolArea, lower = 1)),
+    MiscVal = log(clamp(MiscVal, lower = 1)),
+    SalePrice = log(clamp(SalePrice, lower = 1)),
+  )
+
+# Perform any transformations on the test data as well
+housesTest = housesTest |>
+  mutate(
+    LotArea = log(clamp(LotArea, lower = 1)),
+    FirstFlrSF = log(clamp(`1stFlrSF`, lower = 1)),
+    SecondFlrSF = log(clamp(`2ndFlrSF`, lower = 1)),
+    GrLivArea = log(clamp(GrLivArea, lower = 1)),
+    WoodDeckSF = log(clamp(WoodDeckSF, lower = 1)),
+    OpenPorchSF = log(clamp(OpenPorchSF, lower = 1)),
+    EnclosedPorch = log(clamp(EnclosedPorch, lower = 1)),
+    ThirdSsnPorch = log(clamp(`3SsnPorch`, lower = 1)),
+    ScreenPorch = log(clamp(ScreenPorch, lower = 1)),
+    PoolArea = log(clamp(PoolArea, lower = 1)),
+    MiscVal = log(clamp(MiscVal, lower = 1))
+  )
+```
+
+Define a function to determine the PRESS of each model
+```{r}
+#| eval: false
+#| output: false
+
+PRESS = function(model) {
+    i = residuals(model)/(1 - lm.influence(model)$hat)
+    sum(i^2)
+}
+```
+
+## Analysis 1
+
+Create models for Sales Price for the neighborhood Brookside, Edwards, and North Ames neighborhoods.
+```{r}
+#| eval: false
+#| output: false
+
+housesTrain |>
+  ggplot(aes(x = GrLivArea, y = SalePrice)) +
+  geom_point()
+
+housesTrainNeighborhood = housesTrain |>
+  filter(Neighborhood %in% c("BrkSide", "Edwards", "NAmes"))
+
+housesTrainNeighborhood$Neighborhood = 
+  factor(housesTrainNeighborhood$Neighborhood, levels =
+           c("NAmes", "BrkSide", "Edwards"))
+
+housesTrainNeighborhood |>
+  group_by(Neighborhood) |>
+  summarize(n = n())
+
+housesTrainNeighborhood |>
+  ggplot(aes(x = GrLivArea, y = SalePrice, color = Neighborhood)) +
+  geom_point() +
+  facet_wrap(~Neighborhood) +
+  theme(legend.position="none") +
+  xlab("Log Value of Living Area Square Footage") +
+  ylab("Log Value of Sales Price") +
+  ggtitle("Home Sales Price and Living Area")
+
+model1 = lm(SalePrice ~ Neighborhood, data = housesTrainNeighborhood)
+summary(model1)
+PRESS(model1)
+
+model2 = lm(SalePrice ~ Neighborhood + GrLivArea, 
+            data = housesTrainNeighborhood)
+summary(model2)
+PRESS(model2)
+
+model3 = lm(SalePrice ~ Neighborhood + GrLivArea + Neighborhood:GrLivArea, 
+            data = housesTrainNeighborhood)
+summary(model3)
+PRESS(model3)
+
+plot(model3)
+
+housesTrainNeighborhood |>
+  select(Neighborhood, GrLivArea, SalePrice) |>
+  mutate(SalePricePredicted = predict(model3, 
+                                      newdata = tibble(
+                                        Neighborhood = Neighborhood, 
+                                        GrLivArea = GrLivArea))) |>
+  ggplot() +
+  geom_point(
+    aes(x = exp(GrLivArea), y = exp(SalePrice), color = Neighborhood)) +
+  geom_line(
+    aes(x = exp(GrLivArea), y = exp(SalePricePredicted), 
+        color = Neighborhood)) +
+  facet_wrap(~Neighborhood) +
+  theme(legend.position="none")
+
+```
+
+## Analysis 2
+
+Forward step regression using olss
+
+```{r}
+#| eval: false
+#| output: false
+
+forwardResults = ols_step_forward_aic(
+  lm(SalePrice ~ ., data = housesTrain), details = TRUE, progress = TRUE)
+forwardModel = forwardResults$model
+summary(forwardModel)
+
+# Plot the stepping results impact on AIC
+plot(forwardResults)
+
+# Plot the residual / leverage functions
+plot(forwardModel)
+
+# Calculate the PRESS for the model
+tibble(resids = residuals(forwardModel), 
+       hats = tidy(hatvalues(forwardModel))$x) |>
+  filter(hats != 1) |>
+  mutate(i = (resids / (1-hats))^2) |>
+  select(i) |>
+  pull() |>
+  sum()
+
+# Calculate values against the test data set
+results = housesTest
+results$SalePrice = exp(predict.lm(forwardModel, newdata = housesTest))
+results = results |>
+  dplyr::select(Id, SalePrice)
+
+# Export and submit to Kaggle
+write_csv(results, "Submissions/Forward_Model.csv")
+```
+
+Backward step regression
+
+```{r}
+#| eval: false
+#| output: false
+
+backwardResults = ols_step_backward_aic(
+  lm(SalePrice ~ ., data = housesTrain), details = TRUE, progress = TRUE)
+backwardModel = backwardResults$model
+summary(backwardModel)
+
+# Plot the stepping results impact on AIC
+plot(backwardResults)
+
+# Plot the residual / leverage functions
+plot(backwardModel)
+
+# Calculate the PRESS for the model
+tibble(resids = residuals(backwardModel), 
+       hats = tidy(hatvalues(backwardModel))$x) |>
+  filter(hats != 1) |>
+  mutate(i = (resids / (1-hats))^2) |>
+  select(i) |>
+  pull() |>
+  sum()
+
+# Calculate values against the test data set
+results = housesTest
+results$SalePrice = exp(predict.lm(backwardModel, newdata = housesTest))
+results = results |>
+  dplyr::select(Id, SalePrice)
+
+# Export and submit to Kaggle
+write_csv(results, "Submissions/Backward_Model.csv")
+```
+
+Stepwise regression
+
+```{r}
+#| eval: false
+#| output: false
+
+stepResults = ols_step_both_aic(
+  lm(SalePrice ~ ., data = housesTrain), details = TRUE)
+
+# ols_step_both_aic does not have a direct output for the model determined
+#so we have to write the results as we go and then copy the final model
+#produced and paste it below
+stepModel = lm(SalePrice ~ OverallQual + GrLivArea + Neighborhood + OverallCond
+               + YearBuilt + LotArea + RoofMatl + KitchenAbvGr + SaleCondition
+               + Condition2 + Foundation + Fireplaces + Heating + ExterQual
+               + Condition1 + PoolArea + ScreenPorch + WoodDeckSF + HeatingQC
+               + CentralAir + BedroomAbvGr + FirstFlrSF + `2ndFlrSF` + Street
+               + LandSlope + HalfBath + EnclosedPorch + SecondFlrSF + MiscVal
+               + PavedDrive + BldgType + `1stFlrSF` + ExterCond + LotConfig
+               + YearRemodAdd, data = housesTrain)
+# Plot the stepping results impact on AIC
+plot(stepResults)
+
+# Plot the residual / leverage functions
+plot(stepModel)
+
+# Calculate the PRESS for the model
+tibble(resids = residuals(stepModel), hats = tidy(hatvalues(stepModel))$x) |>
+  filter(hats != 1) |>
+  mutate(i = (resids / (1-hats))^2) |>
+  select(i) |>
+  pull() |>
+  sum()
+
+# Calculate values against the test data set
+results = housesTest
+results$SalePrice = exp(predict.lm(stepModel, newdata = housesTest))
+results = results |>
+  select(Id, SalePrice)
+
+# Export and submit to Kaggle
+write_csv(results, "Submissions/Step_Model.csv")
+```
+
+Custom model using all variables - note, this was unable to achieve higher accuracy than the forward selection results
+
+```{r}
+#| eval: false
+#| output: false
+
+customModel = lm(SalePrice ~ I(MSSubClass^2) + I(LotArea)^2 + Street + LotShape
+                 + LandContour + LotConfig + LandSlope + Neighborhood
+                 + Condition1 + Condition2 + BldgType + HouseStyle
+                 + I(OverallQual^2) + OverallCond + I(YearBuilt^2)
+                 + YearRemodAdd + RoofStyle + RoofMatl + ExterQual + ExterCond
+                 + Foundation + Heating + HeatingQC + CentralAir
+                 + log(`1stFlrSF`) + `2ndFlrSF` + LowQualFinSF + I(GrLivArea^2)
+                 + FullBath + HalfBath + BedroomAbvGr + KitchenAbvGr
+                 + I(TotRmsAbvGrd^2) + I(Fireplaces^2) + PavedDrive
+                 + WoodDeckSF + OpenPorchSF + EnclosedPorch + `3SsnPorch`
+                 + ScreenPorch + PoolArea + MiscVal + MoSold
+                 + YrSold + SaleCondition, data = housesTrain)
+
+summary(customModel)
+
+# Calculate values against the test data setresults = housesTest
+results = housesTest
+results$SalePrice = exp(predict(customModel, newdata = housesTest))
+results = results |>
+  select(Id, SalePrice)
+
+# Export and submit to Kaggle
+write_csv(results, "Submissions/Custom_Model.csv")
+```
